\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-20 notes/2loopNLSM.tex
Small edits.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Comment the following out to compile new images
%
%\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
%\usetikzlibrary{external}  
%\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
%\tikzexternalize[                     %% Activate externalization
% prefix=images/,                    %% Avoid cluttering the directory
% system call={                       %% Use lualatex in system call
%    pdflatex \tikzexternalcheckshellescape -halt-on-error -shell-escape -interaction=batchmode -jobname="\image" "\texsource" || rm "\image.pdf"
% },
%]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\section{One-loop cubic construction}
\label{sec:cubic}
Here we will describe why constructing the full Yang-Mills kinematic algebra (outside self-dual sector) is generically hard, and why the same reason makes NLSM beyond one-loop hard as well. It has to do with how the vector state sum mixes with the four-point contact. Starting with the Yang-Mills Lagrangian,
\begin{equation}
\mathcal{L}^{\text{YM}} = -\frac{1}{2}(\partial_\mu A_\nu)^2 + f^{abc} \partial_\mu A^a_\nu A^b_\mu A^c_\nu + f^{abe}f^{ecd}A^a_\mu A^b_\nu  A^c_\mu A^d_\nu 
\end{equation}
we can express the YM three-point interaction in terms of the following vertex in Lorenz gauge
\begin{equation}
T^k_{ij} = (\varepsilon_i\varepsilon_j)(\varepsilon_k p_i).
\end{equation}
This kinematic generator is anti-symmetric in $1\leftrightarrow 2$ exchange, and from it we can construct the full Lorenz gauge Yang-Mills vertex by summing over cyclic permutations
\begin{equation}
V^{(123)}_{\text{YM}} = T^3_{12} +\text{cyc}(123).
\end{equation}
Contracting this with the vector state sum, we can see that the cubic Yang-Mills vertex does not satisfy the four-point Jacobi identity
\be\label{cubicJac}
{}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4+\text{cyc}(234) = \varepsilon_{(12)}  \varepsilon_{(34)} (s_{13}-s_{23}) +\text{cyc}(234) \neq 0
\end{equation}
where $ \varepsilon_{(ij)} =  \varepsilon_{i}\cdot  \varepsilon_{j}$.
The remaining terms can be absorbed into the definition of the four-point kinematic numerator
\begin{equation}
{}_1\langle V^{2}_{\text{YM}+B}V^{3}_{\text{YM}+B}\rangle_4 = {}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4 + s_{12} (\varepsilon_{(13)}\varepsilon_{(24)}-\varepsilon_{(14)}\varepsilon_{(23)})
\end{equation}
where $V_{\text{YM}+B}$ indicates the introduction of an additional two-form studied in \cite{Ben-Shahar:2022ixa} for constructing the NMHV Yang-Mills Lagrangian.
After the introduction of the 2-form, the numerators now satisfy the Jacobi identity
\be
{}_1\langle V^{2}_{\text{YM}+B}V^{3}_{\text{YM}+B}\rangle_4+\text{cyc}(234) =0
\ee
where the new cubic Lagrangian takes the form
\be
\mathcal{L}^{\text{YM}+B} = -\frac{1}{2}(\partial_\mu A_\nu)^2 - B_{\mu\nu}\Box\tilde{B}_{\mu\nu}+ f^{abc} \partial_\mu A^a_\nu A^b_\mu A^c_\nu+ f^{abc} (B_{\mu\nu}+\Box \tilde{B}_{\mu\nu})^a A^b_\mu A^c_\nu\, .
\ee 
Of course, higher multiplicity would require further redefinition of the three-point vertex to satisfy Jacobi identites on all internal edges. While introducing successively higher spin states could in principle work at tree level, we will see that it is not consistent with what we find at general loop order. We will comment on this in \sect{}. For now, we will study how to construct one-loop integrands consistent with the color-kinematics by isolating the cubic sector of the state sum above.

As can be seen above, the Jacobi identity of \cref{cubicJac} fails only in terms with two factors of polarization dot products, $(\varepsilon\varepsilon)^2$. Whereas, if we restricted ourselves to just considering factors with a single polarization dot product, the Jacobi identity of Yang-Mills would be satisfied \textit{off-shell} and in arbitrary spacetime dimensions
\begin{equation}
{}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4\big|^{(\varepsilon\varepsilon)^1}+\text{cyc}(234) =0\,.
\end{equation}
Thus for our approach at one-loop, we aim to build color-dual integrands directly from kinematic factors of this form. The organizational principle underlying this construction can be understood in terms of a polarization decomposition of vector amplitudes introduced by one of the authors \cite{Pavao:2022kog}, 
\begin{equation}\label{eq:pureVecRABD}
A_{(\sigma)}^{\text{YM}} = \sum_{k=0}^{\lfloor |\sigma|/2\rfloor}\sum_{\rho \in S^{2|k}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\end{equation}
\jm{Nic, could you give a quick recap of what all of the symbols mean in the equation above?}
By a simple mass-dimensioin argument, we know the tree-level cubic sector of Yang-Mills must be at $\mathcal{O}((\varepsilon\varepsilon)^1)$,
\begin{equation}
A_{(\text{tree})}^{\text{cubic}} = \sum_{\rho \in S^{2|1}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\, ,
\end{equation}
while at one-loop, the cubic sector corresponds to $\mathcal{O}((\varepsilon\varepsilon)^0)$ in polarization dot products,
\begin{equation}
A_{(\text{1-loop})}^{\text{cubic}} = \sum_{\rho \in S^{2|0}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\end{equation}
When decomposed in this way, the polarization stripped building blocks must obey a set of Ward-identities between different ``helicity'' (or $(\varepsilon \varepsilon)$) sectors in order for the full amplitude to be gauge invariant,
 \begin{equation}\label{eq:GIrelA}
\Delta_{(\sigma)}^{(\rho)}\Big|_{\epsilon_i\rightarrow k_i} =-
\sum_{j \in \rho^c} (k_i \varepsilon_j)\Delta_{(\sigma )}^{(\rho\cup (ij))} \,.
\end{equation}
With this in hand, we can construct a Lagrangian description of the of the cubic sector for Yang-Mills, and will demonstrate that the resulting amplitudes with manifestly color-dual Feynman rules are equivalent to both SDYM and NLSM through one-loop. 
\subsection{Semi-abelian Yang-Mills theory}\label{semiYM}
As we demonstrate above, if we include factors of $(\varepsilon\varepsilon)^{n\geq 1}$ at tree-level, then we need to keep the Yang-Mills four-point contact for color-kinematics duality to be restored on-shell. However, the contrapositive is also true -- if we omit the four-point Yang-Mills vertex, then we only need terms that contribute to the manifestly color-dual cubic sector of the theory.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Comment the following out to compile new images

\usetikzlibrary{external}  
\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\tikzexternalize[                     %% Activate externalization
prefix=images/,                    %% Avoid cluttering the directory
system call={                       %% Use lualatex in system call
   pdflatex \tikzexternalcheckshellescape -halt-on-error -shell-escape -interaction=batchmode -jobname="\image" "\texsource" || rm "\image.pdf"
},
]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\section{One-loop cubic construction}
\label{sec:cubic}
Here we will describe why constructing the full Yang-Mills kinematic
algebra (outside the self-dual sector) is generically hard, and why
the same reason makes NLSM beyond one-loop hard as well. The major
obstacle stems from the vector state sum mixing with higher-point
contacts. Starting with the Yang-Mills Lagrangian,
\begin{equation}
  \mathcal{L}^{\text{YM}} = -\frac{1}{2}(\partial_\mu A_\nu)^2
  + f^{abc} \partial_\mu A^a_\nu A^b_\mu A^c_\nu
  + f^{abe}f^{ecd}A^a_\mu A^b_\nu  A^c_\mu A^d_\nu 
\end{equation}
we can express the YM three-point interaction in terms of the
following vertex in Lorenz gauge
\begin{equation}
T^k_{ij} = (\varepsilon_i\varepsilon_j)(\varepsilon_k p_i).
\end{equation}
This kinematic generator is anti-symmetric in $1\leftrightarrow 2$
exchange, and from it we can construct the full Lorenz gauge
Yang-Mills vertex by summing over cyclic permutations
\begin{equation}
V^{(123)}_{\text{YM}} = T^3_{12} +\text{cyc}(123).
\end{equation}
Contracting this with the vector state sum, we can see that the cubic
Yang-Mills vertex does not satisfy the four-point Jacobi identity
\begin{equation}
  \label{cubicJac}
  {}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4+\text{cyc}(234) = \varepsilon_{(12)}  \varepsilon_{(34)} (s_{13}-s_{23}) +\text{cyc}(234) \neq 0
\end{equation}
where $ \varepsilon_{(ij)} = \varepsilon_{i}\cdot \varepsilon_{j}$.
The remaining terms can be absorbed into the definition of the
four-point kinematic numerator
\begin{equation}
{}_1\langle V^{2}_{\text{YM}+B}V^{3}_{\text{YM}+B}\rangle_4 = {}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4 + s_{12} (\varepsilon_{(13)}\varepsilon_{(24)}-\varepsilon_{(14)}\varepsilon_{(23)})
\end{equation}
where $V_{\text{YM}+B}$ indicates the introduction of an additional
two-form studied in Ref.~\cite{Ben-Shahar:2022ixa} for constructing the
NMHV Yang-Mills Lagrangian.  After the introduction of the 2-form, the
numerators now satisfy the Jacobi identity \be {}_1\langle
V^{2}_{\text{YM}+B}V^{3}_{\text{YM}+B}\rangle_4+\text{cyc}(234) =0 \ee
where the new cubic Lagrangian takes the form \be
\mathcal{L}^{\text{YM}+B} = -\frac{1}{2}(\partial_\mu A_\nu)^2 -
B_{\mu\nu}\Box\tilde{B}_{\mu\nu}+ f^{abc} \partial_\mu A^a_\nu A^b_\mu
A^c_\nu+ f^{abc} (B_{\mu\nu}+\Box \tilde{B}_{\mu\nu})^a A^b_\mu
A^c_\nu\, .  \ee Of course, higher multiplicity would require further
redefinition of the three-point vertex to satisfy Jacobi identites on
all internal edges. While introducing successively higher spin states
could in principle work at tree level, we will see that it is not
consistent with what we find at general loop order. We will comment on
this in \sect{}. For now, we will study how to construct one-loop
integrands consistent with the color-kinematics by isolating the cubic
sector of the state sum above.

As can be seen above, the Jacobi identity of \cref{cubicJac} fails
only in terms with two factors of polarization dot products,
$(\varepsilon\varepsilon)^2$. Whereas, if we restricted ourselves to
just considering factors with a single polarization dot product, the
Jacobi identity of Yang-Mills would be satisfied \textit{off-shell}
and in arbitrary spacetime dimensions
\begin{equation}
{}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4\big|^{(\varepsilon\varepsilon)^1}+\text{cyc}(234) =0\,.
\end{equation}
Thus for our approach at one-loop, we aim to build color-dual
integrands directly from kinematic factors of this form. The
organizational principle underlying this construction can be
understood in terms of a polarization decomposition of vector
amplitudes introduced by one of the authors \cite{Pavao:2022kog},
\begin{equation}\label{eq:pureVecRABD}
A_{(\sigma)}^{\text{YM}} = \sum_{k=0}^{\lfloor |\sigma|/2\rfloor}\sum_{\rho \in S^{2|k}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\end{equation}
\jm{Nic, could you give a quick recap of what all of the symbols mean
  in the equation above?}  By a simple mass-dimensioin argument, we
know the tree-level cubic sector of Yang-Mills must be at
$\mathcal{O}((\varepsilon\varepsilon)^1)$,
\begin{equation}
A_{(\text{tree})}^{\text{cubic}} = \sum_{\rho \in S^{2|1}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\, ,
\end{equation}
while at one-loop, the cubic sector corresponds to
$\mathcal{O}((\varepsilon\varepsilon)^0)$ in polarization dot
products,
\begin{equation}
A_{(\text{1-loop})}^{\text{cubic}} = \sum_{\rho \in S^{2|0}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\end{equation}
When decomposed in this way, the polarization stripped building blocks
must obey a set of Ward-identities between different ``helicity'' (or
$(\varepsilon \varepsilon)$) sectors in order for the full amplitude
to be gauge invariant,
 \begin{equation}\label{eq:GIrelA}
\Delta_{(\sigma)}^{(\rho)}\Big|_{\epsilon_i\rightarrow k_i} =-
\sum_{j \in \rho^c} (k_i \varepsilon_j)\Delta_{(\sigma )}^{(\rho\cup (ij))} \,.
\end{equation}
With this in hand, we can construct a Lagrangian description of the of
the cubic sector for Yang-Mills, and will demonstrate that the
resulting amplitudes with manifestly color-dual Feynman rules are
equivalent to both SDYM and NLSM through one-loop.

\subsection{Semi-abelian Yang-Mills theory}\label{semiYM}
As we demonstrate above, if we include factors of
$(\varepsilon\varepsilon)^{n\geq 1}$ at tree-level, then we need to
keep the Yang-Mills four-point contact for color-kinematics duality to
be restored on-shell. However, the contrapositive is also true -- if
we omit the four-point Yang-Mills vertex, then we only need terms that
contribute to the manifestly color-dual cubic sector of the theory.
